Who Needs Photo-z? A Variational Approach to Inferring the Redshift Distribution of Weak Lensing Source Galaxies in 3x2pt Analyses Without Photometric Redshifts

### 2022-12-12 src/tex/ms.tex
Adding new text

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

@@

\section{Method}

\subsection{3x2pt Analysis framework}

@@

\subsubsection{3x2pt Systematics Modeling}

@@

>>>>

% Define document class
\documentclass[twocolumn,draft]{aastex631}
\usepackage{showyourwork}

@@

The paper is organized as follows. In Section 2, we describe the weak lensing observables and the 3x2pt analysis method used in this work. In Section 3, we introduce our approach for inferring the redshift distribution of weak lensing source galaxies using variational inference. In Section 4, we present the results of our analysis on simulated DES Y1 data, and compare our results to those obtained using photometric redshifts. Finally, in Section 5, we discuss the implications of our results and conclude.

\section{Weak Lensing and the 3x2pt Analysis}

@@

\subsection{3x2pt Systematics Modeling}

@@

\section{Inferring Redshift Distributions with Variational Inference}

\subsection{Introduction to variational inference}

Variational inference is a powerful tool for estimating complex distributions from data. It is a form of approximate Bayesian inference, which seeks to find a tractable distribution that approximates the true posterior distribution of the model parameters. This is typically done by minimizing the Kullback-Leibler divergence between the approximate distribution and the true posterior (Blei et al., 2017).

In the context of our analysis, we are interested in inferring the redshift distribution of weak lensing source galaxies from the data. This is a complex problem, as the redshift distribution is a high-dimensional function that depends on both the cosmological parameters, $\theta_{cosmo}$, and the parameters describing the redshift distributions, $\theta_{nzs}$. We can use variational inference to find an approximate distribution $q(\theta_{cosmo},\theta_{nzs})$ that approximates the true posterior distribution $p(\theta_{cosmo},\theta_{nzs}|d)$, where $d$ is the data.

To do this, we define the variational objective, also known as the evidence lower bound (ELBO), as follows:
\begin{equation}
\text{ELBO} = \mathbb{E}_{q(\theta_{cosmo},\theta_{nzs})} \left[ \log p(d|\theta_{cosmo},\theta_{nzs}) \right] - \text{KL}\left[ q(\theta_{cosmo},\theta_{nzs}) | p(\theta_{cosmo},\theta_{nzs})\right]
\end{equation}
where $\mathbb{E}{q(\theta{cosmo},\theta_{nzs})}$ denotes the expectation with respect to the approximate distribution $q(\theta_{cosmo},\theta_{nzs})$, $\log $

\subsection{Parametric form for the redshift distribution}

One of the key challenges in modeling the redshift distribution of weak lensing source galaxies is finding a flexible and accurate parametric form for the distribution. In this analysis, we propose using a mixture of Beta distributions to model the redshift distribution for each tomographic bin.

Beta distributions are a family of continuous probability distributions defined on the interval [0,1]. They have two shape parameters, which control the location and shape of the distribution, and are commonly used to model proportions and probabilities. Mixtures of Beta distributions allow for more flexibility and modeling power than using a single Beta distribution, and can accurately capture a wide range of redshift distributions.

The redshift distribution for each tomographic bin is modeled as a mixture of K Beta distributions:

\begin{equation}
p(z) = \frac{1}{z_{max}} \sum_{k=1}^{K} w_k \cdot Beta\left(\frac{z}{z_{max}}; \alpha_k, \beta_k\right)
\end{equation}

where $w_k$ is the weight of the kth component, and $\alpha_k$ and $\beta_k$ are the shape parameters of the kth Beta distribution. The weights are normalized such that $\sum_{k=1}^{K} w_k = 1$. $z_{max}$ is the maximum redshift of the distribution, which is typically set to the maximum observed redshift in the data.

The mixture of Beta distributions model offers several advantages over other commonly used parametric forms in cosmology. For example, compared to Gaussian distributions, the mixture of Beta distributions model is more flexible and can capture a wider range of shapes, including multi-modal distributions. It is also bounded at 0, which is important for modeling the redshift distribution of weak lensing source galaxies.
